Artificial ecosystem selection reveals relationships between microbiome composition and ecosystem function

### 2022-10-19 Manuscript/manuscript.Rmd
update everything for manuscript draft to lab group

Microbiomes mediate important ecosystem functions, yet it has proven difficult
to determine the relationship between microbiome composition and the rate of
any ecosystem function. This challenge remains because it is difficult to
manipulate microbiome composition directly, we cannot know a priori which members of the
microbiome limit the rate of an ecosystem function, and microbiomes covary with
the environment. We propose that artificial ecosystem selection
can overcome these challenges. Artificial ecosystem selection involves creating
replicate ecosystems, selecting replicates with a desired ecosystem-level
trait, and using the microbiomes from the selected ecosystems to inoculate a
new set of ecosystems. We performed artificial ecosystem selection on replicate
soil microcosms to select for increased soil methane oxidation rate. We
observed a strong response to selection with a 51% increase in methane oxidation rate per round over five
rounds of selection. This suggests that a large portion of the variation in
methane oxidation rate in these soils can be attributed to microbiome
variation. We also found that selection did not enrich for known methane
oxidizers; rather microbes that responded to selection included `r length(da_fams)`
families such as *Fimbriimonadaceae*, *Cytophagaceae*, and *Diplorickettsiaceae*. This result
is in contrast to the typical assumption that the rate of an ecosystem function
is limited by the final step in the associated microbial pathway. Our study
demonstrates that variation in microbiome composition can contribute to
variation in the rate of ecosystem function.

@@

Microbiomes mediate a variety of important functions in ecosystems, and there
is great interest in understanding how attributes of microbiomes may influence
variation in ecosystem functions [@crowther2019]. Biodiversity-ecosystem
function relationships have been described for a variety of macroorganismal
communities. For example, plant species richness correlates with productivity
and marine community diversity correlates with functions such as production,
consumption, and nutrient release
[@cardinale2006; @hooper2012; @gamfeldt2015]. While there is evidence that
variation in the microbiome contributes to variation in ecosystem functions,
describing this relationship has been elusive [@rocca2015; @graham2016].

One reason for this difficulty is that it is challenging to directly manipulate
microbiome composition and diversity the way we can for macroorganismal
communities. Instead, we must use genetics through marker gene or metagenomic
sequencing to characterize the microbiome.  However, studies that measure the
taxonomic or functional diversity of a microbiome through these approaches
rarely find a correlation between diversity and function [@graham2016]. Another
approach is to quantify cells or genes using  quantitative
PCR. Studies such as this still rarely find a correlation between the abundance
of a functional group and the rate of the corresponding process [@rocca2015].
As a result, it remains unclear how and when microbiome variation contributes
to variation in the rate of ecosystem functions.

Another likely reason that microbiome-ecosystem function relationships have been
difficult to document is that the most common approaches for looking for a
relationship requires some prior knowledge of the likely causal links between
microbiomes and ecosystem functions. For example, many studies have looked for
relationships with the abundance or diversity of a "marker gene" for a
particular function. Marker genes are microbial genes that code for an enzyme or
enzyme-subunit known to be involved in a particular function. But this approach
assumes that one can identify the "right" marker gene a priori which, given
how little is understood about microbial diversity, seems highly unlikely. What
is needed are approaches to microbiome-ecosystem
function mapping that do not require this degree of prior knowledge.

Yet another likely reason that microbiome-ecosystem function relationships have
been difficult to document is that microbiomes can be related to ecosystem
functions in two related but distinct ways that historically have been difficult
to separate. One way is that microbiomes may simply be conduits through which
the environment drives function. That is to say that the environmental
conditions completely regulate the attributes of the microbiome and a shift in
the environment causes a shift in microbiome attributes, such as the abundance
of a microbial functional group. This shift results in a change in the rate of
an ecosystem function. In this case, microbiome attributes would be
statistically associated with variation in ecosystem function, but also strongly
covary with environmental conditions. Therefore, there would be no association
between microbiomes and ecosystem functions independent of the environmental
conditions. Under this scenario, knowing anything about the microbiome would not
necessarily provide you with information regarding the causes of the change in
ecosystem function or improve one's ability to predict changes in ecosystem
function. Furthermore, attempting to alter ecosystem function by altering the
microbiome without changing the underlying environmental conditions would likely
be fruitless.

Another possibility is that the microbiome itself alters the rate of ecosystem
function independent of the environment. In this case, one cannot predict the
rate of ecosystem function without understanding changes in the microbiome. In
addition, managing the microbiome through inoculation or selective antibiotics
would potentially be an effective approach for enhancing or mitigating certain
changes to ecosystem functions. These two scenarios are not mutually
exclusive, nor are they likely unique to microbes, but methods to determine the
relative importance of these two scenarios have not until recently been employed.

The problem of connecting microbiome attributes to the rate of an ecosystem
function is analogous to the problem of connecting genomic variation to
phenotypic traits in organisms, and solutions to this problem may be found by
exploring this analogy [@morris2020]. For example, one approach to the problem of
genotype-phenotype mapping is to perform artificial selection on a trait of
interest and identify genes that respond to selection. Such genes are likely to
be causally linked to the trait under selection. An analogous approach for
exploring microbiome-ecosystem function relationships is gaining acceptance
among microbiologists [@swenson2000; @panke-buisse2015]. Microbiomes have been
shown to respond to selection at the ecosystem-trait level [@swenson2000],
and some ecosystem-scale microbiome traits have been shown to be transferable
through microbiome inoculations [@panke-buisse2015]. Applying artificial
selection to whole microbiomes could be a novel approach for quantifying how
much of the variation in the rate of an ecosystem function could be attributed
independently to microbiome variation [@goodnight2000]. In addition, this
approach could be used to identify microbiome attributes, such as genes or taxa,
associated with the rate of an ecosystem function, although, to our knowledge
this has not previously been accomplished [@arias-sanchez2019; @morris2020]. 
Such an approach has the potential to expand our ability to model
microbial ecosystem-function relationships more accurately and allow us to
manage ecosystems for particular outcomes [@fuhrman2009; @schimel1998]. Here, we
use this approach to estimate how much of the variation in ecosystem function
(the flux of CH~4~ gas between the soil and atmosphere) can be attributed to
variation in the microbiome and to identify microbial markers associated with
this ecosystem function.

Soil CH~4~ oxidation is a suitable function for this study because there is
evidence that it may vary with microbiome attributes. For example, variation in
CH~4~ emissions in arctic permafrost is correlated with the transcriptional
activity of certain methanogens [@freitag2009; @freitag2010]. In addition,
soil microbiome phylogenetic variation is a strong predictor of CH~4~ oxidation in
forests and pastures of the Brazilian Amazon [@meyer2020]. Finally,
methanogenesis and methanotrophy are two of the most deeply conserved microbial
physiologies and are represented in a narrow range of taxa and so the taxonomic
composition of the microbiome is more likely to be associated with the rate of
CH~4~ flux than other broader or more shallowly conserved functions
[@schimel1998; @martiny2013]. This suggests that microbiome variation might
drive variation in CH~4~ oxidation in soils.

We applied artificial ecosystem selection on soil microbiomes by selecting on
the ecosystem-scale CH~4~ oxidation rate.  Using this approach, we were able to
estimate how much of the variation in CH~4~ oxidation rate was associated with
variation in the microbiome.  To identify which microorganisms could be markers
of CH~4~ oxidation rate, we compared the composition of the artificially
selected microbiomes to a control set of microbiomes without selection.  We
then evaluated whether these markers met our underlying assumptions about
which taxa limit the rate of ecosystem function.

@@

Methane oxidation rates were determined after flushing and spiking jars to
approximately 1000 ppm CH~4~. Headspace samples of 1 mL were collected from each
jar immediately after spiking and then at time points 3, 6, 24, and 48 hours for
a 5-point curve. Samples were immediately injected into a SRI model 8610C gas
chromatograph equipped with a flame ionization detector (SRI Instruments, Torrance, CA, USA) to determine the headspace CH~4~
concentration. Fluxes were calculated from a first-order exponential decay
function as decay constant $k$ with units day^-1^. Oxidation rates are presented
as the additive inverse of $k$ (i.e., $-k$) so that a more positive value
represents a greater oxidation rate. The jars selected for the positive
treatment in passage 2 had the lowest CH~4~ oxidation rate of the twelve jars
due to a calculation error in the CH~4~ oxidation rate. All other passages
correctly used the top three jars.

@@

Bioinformatics processing was performed in 'R' [@rcoreteam2018].
Demultiplexed sequencing reads were denoised using 'DADA2' to generate a table
of amplicon sequence variants (ASVs)
[@callahan2016]. Taxonomic assignment was performed
using the Ribosomal Database Project naive Bayesian classifier [@wang2007]. The presence of
contaminants was evaluated using both the prevalence and frequency methods from
'DECONTAM' [@davis2018].
Decontam identified `r length(which(contamdf$contaminant))` potential contaminants based
on prevalence and frequency. Visual inspection of abundance-concentration
plots indicated that `r length(contaminants)` of these were likely contaminants and these
ASVs were removed.

@@

To estimate the proportion of variance in CH~4~ oxidation rate due to variation in
the microbiome, we estimated narrow-sense heritability ($h^2$) as the
regression of mid-offspring on mid-parent [@falconer1996]. The mid-parent was the
mean for all three selected jars and the mid-offspring was the mean for all jars
produced by those parents. First, we tested if there was an effect of treatment
on the heritability estimate. We compared nested models with and without treatment using an F-test
test with the 'anova' function. We then fit the full model to estimate the
heritability of the Positive and Neutral treatments.

@@

To identify taxa that responded to selection on CH~4~ oxidation rate, we tested
differential abundance between the two treatments in passage 5. We first
grouped ASVs at the family level keeping ASVs that lacked a family-level taxonomic
assignment by grouping them at a higher level. We
then subset the samples in Passage 5 and removed all families with a prevalence
of less than 10\% in either treatment. We used three methods for testing differential abundance: ANCOM-II, ALDEx2, and CORNCOB. We then
pulled out the consensus taxa that were significant
with all three tests and plotted their relative abundances. For ANCOM-II, we used the 'ancom' function in the 'ANCOM-BC' package with a cutoff of W = 0.7 [@kaul2017; @lin2020]. For ALDEx2, we used the 'aldex' function in the 'ALDEx2' package with Welch's t-test and we used an effect size of
1 as our significance threshold [@fernandes2014]. Finally, we used CORNCOB with the 'differentialTest' function in the 'corncob' package with the Wald test and without bootstrapping [@martin2020]. Lastly, to test differentially abundant methanotrophs, we subset all ASVs within methanotrophic families and tested their differential abundance aggregated at the family and genus level using 'corncob'.

@@

We observed a response to selection on whole-ecosystem soil CH~4~ oxidation
rate (`r figures("response", display = "cite")`; difference of slopes:
F~`r compare_response_mods[2,3]`,`r compare_response_mods[2,1]`~ = `r myround(compare_response_mods[2,5], 2)`, p = `r myround(compare_response_mods[2,6], 2)`). At the start of the experiment, the Positive
treatment had a mean CH~4~ oxidation rate that was
`r percent_less(initial = tidy(response_model)[[1, 2]], final = tidy(response_model)[[3, 2]] + tidy(response_model)[[1, 2]])`\% lower than the
Neutral treatment `r print_lm(response_model, 3, 'difference of y-intercepts')`. 
There was no change in CH~4~ oxidation rate of the Neutral
treatment over the five passages `r print_lm(response_model, 2)`. By contrast,
the Positive treatment had a `r round(per_slope)`\% increase in CH~4~
oxidation rate per passage `r print_lm(response_model, 4)`.

## Heritability of methane oxidation

We estimated heritability as the regression of mid-offspring on mid-parent (`r figures("herit", display = "cite")`).
Offspring CH~4~ oxidation rates were correlated with parental CH~4~ oxidation
rates in both the Positive treatment `r print_lm(herit_model, 4)` and the
Neutral treatment `r print_lm(herit_model, 2)`.

@@

# New Discussion

We observed a response to selection on whole-ecosystem soil methane oxidation rate, which demonstrates that variation in the microbiome can generate variation in the rate of ecosystem function independent of the underlying environmental conditions. While it is well understood that the microbiome mediates a variety of ecosystem functions, a fundamental question in microbial ecology is whether we need to understand variation in the composition of microbiomes to understand variation in the rate of ecosystem functions. To address that question, we performed artificial ecosystem selection on soil microbiomes by selecting for ecosystems with high methane oxidation rates. We found that there was a significant increase in methane oxidation rate as a response to selection. Each generation of soil ecosystems was created from the same pool of sterile soil substrate and a small inoculum from the previous generation. Therefore, it is unlikely that this response to selection was due to changes in the abiotic conditions of the ecosystem. In addition, the response was only observed in the Positive treatment and not in the Neutral treatment, which shows that this was not a side effect of passaging in jars. Therefore, we conclude that variation in the microbiome can generate variation in ecosystem functions independent of the environment.

In addition, we observed a large heritability for methane oxidation rate, which suggests that microbiome variation explains a large portion of the variation in methane oxidation rate relative to environmental variation in this system.
To estimate how much of the variation in methane oxidation rate can be explained by the presence, absence, and relative abundances of specific microbial taxa, we calculated the narrow-sense heritability of methane oxidation rate, which was `r myround(coefficients(herit_model)[4], 2)` for the Positive treatment. A heritability of 1 would indicate that a trait is entirely genetically determined and thus offspring would have the exact same value of that trait as their parents. A heritability greater than 1 indicates that offspring exceed the value of their parents. This is unlikely in the context of a genomic trait, but microbiomes are not limited by the constraints of genomes because microbes are not alleles and communities are not chromosomes. Therefore, complementarity within a microbial community could result in offspring communities that exceed the trait values of their "parent" ecosystems. The large heritability we observed indicates that microbiome variation is a large contributor to the variation in methane oxidation in this experimental system.

@@

Richness at the ASV level did not vary between the two treatments and there were relatively few taxa gained or lost in the Positive selection treatment and none of these were prevalent across the 12 jars in passage 5. Therefore, the gain or loss of species is unlikely to explain the increase in methane oxidation rate. However, we found that Bray-Curtis dissimilarity was greater between the two treatments in Passage 5 than within each treatment and was correlated with methane oxidation rate, which suggests that changes in the relative abundance of taxa could explain the response to selection. A difference in composition between the two treatments could be a direct result of selection for organisms that contribute to CH~4~ oxidation or a result of ecological drift over the course of the experiment. In the former case, that would suggest that changes in the relative abundance of specific microbial taxa could explain the response to selection on CH~4~ oxidation rate.

Even though we observed an increase in methane oxidation rate in the Positive treatment and a difference in composition between the two treatments, we did not observe an increase in the relative abundance of methanotrophs. This contrasts with the usual assumption that the final step in a metabolic pathway is the rate-limiting step (citation?). In certain ecosystems, methane oxidation rate is correlated with the abundance of methanotrophs as estimated from marker genes (Cite. I know that one methanogenesis paper, not sure about methanotrophy). However, it does not appear generally true that the rate of an ecosystem function is limited by the abundance of the gene that encodes the final step in that pathway (cite rocca, maybe some of the methane work.) Since we did not observe an increase in the relative abundance of methanotrophs in the treatment that displayed an increase in methane oxidation rate, we may have to consider other microbial functional groups within an ecosystem to understand what drives changes in the rate of ecosystem functions.

One of the goals of this study was to determine how much of the variation in ecosystem function is due to variation in the microbiome. In this experiment, we eliminated as much environmental variation as possible. This was intentional, because we wanted to determine whether it was possible for variation in the microbiome to generate variation in methane oxidation rate independent of the environment. As a result, our heritability estimates are likely to be greater than they would be in natural ecosystems with greater environmental variation. Future work could estimate heritability in experimental systems that manipulate environmental variation or across natural ecosystems to evaluate how much of the variation in ecosystem function can be explained by microbiome variation outside the lab. This experiment also had a relatively small sample size which resulted in heritability estimates with wide error bars. Increasing replication would help us more accurately estimate this parameter and give us a better understanding of the relative importance of environmental variation and microbiome variation in the rate of microbiome-mediated ecosystem functions.

Another important question is whether selection at the ecosystem scale resulted in within-strain evolution. Using 16S rRNA gene sequencing we are unable to resolve genomic changes within taxa that may have undergone evolution. Future work could analyze an artificial ecosystem selection experiment using genome-resolved metagenomics to try to detect strain-level changes in the constituent taxa. This would help evaluate whether certain metabolic pathways were enriched in the positive selection treatment or if functional genes changed in frequency within the microbiome.

As more is understood about the role of the microbiome in regulating ecosystem functions and the specific taxa involved, more detailed experiments could selectively add or remove important taxa or manipulate their relative abundance to start to estimate the quantitative relationship between specific taxa and the rate of an ecosystem function. Both our differential abundance analysis and this reductionist approach to manipulating individual taxa assumes that taxa act individually on ecosystem function. However, it is much more likely that the microbiome affects the rate of ecosystem function through the interactions of multiple taxa and future analyses with greater replication should test the role of multiple taxa simultaneously by, for example, using machine learning methods [@topçuoğlu2020]

A fundamental question in microbial ecology is whether variation in the microbiome contributes to variation in ecosystem function. We performed artificial ecosystem selection on soil methane oxidation rate to test the relationship between microbiome variation and soil methane oxidation rate. We observed a response to selection on soil methane oxidation rate and found a high heritability for methane oxidation, which suggests that variation in the microbiome contributes to variation in soil methane oxidation rate. Surprisingly, we did not observe an increase in the relative of abundance of methanotrophs as a response to selection. Future research could estimate heritability under greater environmental variation to understand how much of the variation in methane oxidation rate is due to variation in the microbiome in a natural ecosystem. In addition, future selection experiments could use genome-resolved metagenomics to observe within-strain evolution and test changes in the functional pathways of the microbiome as a response to selection to determine what aspects of the microbiome explain this increase in oxidation rate. This study demonstrates that soil methane oxidation rate can vary greatly with variation in the microbiome and opens up future research opportunities to establish a more detailed understanding of the relationship between the microbiome and ecosystem function.

# Old Discussion

We observed a response to
selection on whole-ecosystem soil CH~4~ oxidation rate. This type of response
has been observed for other functions, such as chloroaniline degradation in
water [@swenson2000], but this is the first instance, to our knowledge, of
selection being performed directly on soil biogeochemical cycling. With this
experimental design, we attempted to create multiple ecosystems that were very
similar with respect to the type and quantity of soil, soil moisture content,
and the headspace concentration of CH~4~. we further controlled for environmental
variation across replicates by having twelve replicates per treatment, using
controlled laboratory conditions, and by having a random selection treatment as
a control. Therefore, the observed response to selection on
CH~4~ oxidation rate is likely due to changes in the microbiome.

>>>>

Microbiomes mediate important ecosystem functions, yet it has proven difficult
to determine the relationship between microbiome composition and the rate of
ecosystem functions. This challenge remains because it is difficult to
manipulate microbiome composition directly, we often cannot know a priori which
members of a microbiome influence the rate of an ecosystem function, and
microbiomes can covary strongly with other drivers of ecosystem function, such
as the environment. We propose that artificial ecosystem selection can overcome
these challenges. Artificial ecosystem selection involves creating replicate
ecosystems, selecting replicates with a desired ecosystem-level trait, and
using the microbiomes from the selected ecosystems to inoculate a new set of
ecosystems. We performed artificial ecosystem selection on replicate soil
microcosms to select for increased soil methane oxidation rate. We observed a
strong response to selection with a 51% increase in methane oxidation rate per
round over five rounds of selection. This suggests that a large portion of the
variation in methane oxidation rate in these soils can be attributed to
microbiome variation. We also found that selection did not enrich for known
methane oxidizers; rather microbes that responded to selection included members
of `r length(da_fams)` families such as *Fimbriimonadaceae*, *Cytophagaceae*,
and *Diplorickettsiaceae*. This result is in contrast to the typical assumption
that the rate of an ecosystem function is limited by the final step in the
associated microbial pathway. Our study demonstrates that variation in
microbiome composition can contribute to variation in the rate of ecosystem
function and that this relationship can be revealed via artificial ecosystem
selection.

@@

Microbiomes mediate a variety of important functions in ecosystems, and there
is great interest in understanding how attributes of microbiomes may influence
variation in ecosystem functions [@conrad1996; @schimel1998; @crowther2019]. Biodiversity-ecosystem
function relationships have been described for a variety of macroorganismal
communities. For example, plant species richness correlates with productivity
and marine community diversity correlates with functions such as production,
consumption, and nutrient release
[@cardinale2006; @hooper2012; @gamfeldt2015]. While there is some evidence that
variation in microbiomes can contribute to variation in ecosystem functions,
documenting this relationship has been elusive [@conrad1996; @rocca2015; @graham2016].

There are several reasons for this difficulty. One reason is that it is
challenging to manipulate microbiome composition and diversity directly.
Experimental manipulations of community composition and diversity have been
central to documenting the relationships between ecosystem functions and the
biodiversity of plants and animals [@hooper2012], but this is seldom possible
with microbiomes. Most microorganisms cannot yet be cultured and most
available methods for manipulation of microbiomes (e.g., antibiotics, dilution,
etc.) are not sufficiently specific to manipulate individual microbial
lineages. Microbial ecologists are thus often restricted to using comparative
methods, in which microbiome composition and diversity is surveyed via
environmental DNA and then correlated with the rate of an ecosystem function
{cite examples}. However, such studies are rarely successful [@graham2016]. In an
attempt to increase the likelihood of detecting such a correlation, microbial
ecologists may focus their survey efforts on a particular group of microbes
thought to be involved in the ecosystem function of interest. For example, many
studies have looked for relationships between an ecosystem function and the
abundance or diversity of a microbial gene that codes for an enzyme or
enzyme-subunit known to be involved in a particular function. But this approach
assumes that one can identify the “right” marker gene a priori which, given how
little is understood about microbial diversity, seems highly unlikely. And in
fact these studies are rarely successful [@rocca2015].

Another likely reason that microbiome-ecosystem function relationships have
been difficult to document is that microbiomes often covary with other drivers
of variation in ecosystem functions, such as environmental conditions or
resources. Such covariation can make it impossible to determine whether
microbiomes directly contribute to functional variation. Understanding the
degree to which microbiomes directly influence functional variation is crucial
if scientists are to develop microbiome-focused methods for altering or
maintaining ecosystem functions.

The problem of connecting microbiome attributes to the rate of an ecosystem
function is analogous to the problem of connecting genomic variation to
phenotypic traits in organisms, and solutions to this problem may be found by
exploring this analogy [@morris2020]. For example, one approach to the problem
of genotype-phenotype mapping is to perform artificial selection on a trait of
interest and identify genes that respond to selection. Such genes are likely to
be causally linked to the trait under selection. An analogous approach for
exploring microbiome-ecosystem function relationships is gaining acceptance
among microbiologists [@swenson2000; @panke-buisse2015; @sanchez2021].
Microbiomes have been shown to respond to selection at the ecosystem-trait
level [@swenson2000], and some ecosystem-scale microbiome traits have been
shown to be transferable through microbiome inoculations [@panke-buisse2015].
Applying artificial selection to whole microbiomes could be a novel approach
for quantifying how much of the variation in the rate of an ecosystem function
could be attributed independently to microbiome variation [@goodnight2000]. In
addition, this approach could be used to identify microbiome attributes, such
as genes or taxa, associated with the rate of an ecosystem function, although,
to our knowledge this has not previously been accomplished [@arias-sanchez2019;
@morris2020]. Such an approach has the potential to expand our ability to more
accurately model microbial ecosystem-function relationships and allow us to
manage ecosystems for particular outcomes [@fuhrman2009; @schimel1998].

We use this approach to estimate how much of the variation in an ecosystem
function - the flux of CH~4~ gas between the soil and atmosphere - can be
attributed to variation in the microbiome and to identify microbial markers
associated with this ecosystem function.  Soil CH~4~ oxidation is a suitable
function for this study because there is evidence that it may vary with
microbiome attributes. For example, variation in CH~4~ emissions in arctic
permafrost is correlated with the transcriptional activity of certain
methanogens [@freitag2009; @freitag2010]. In addition, soil microbiome
phylogenetic variation is a strong predictor of CH~4~ oxidation in forests and
pastures of the Brazilian Amazon and the Congo Basin [@meyer2020; @meyer2020a]. Finally, methanogenesis and
methanotrophy are two of the most deeply conserved microbial physiologies and
are represented in a narrow range of taxa and so the taxonomic composition of
the microbiome is more likely to be associated with the rate of CH~4~ flux than
other broader or more shallowly conserved functions [@schimel1998;
@martiny2013]. This suggests that microbiome variation might drive variation in
CH~4~ oxidation in soils.

We applied artificial ecosystem selection on soil microbiomes by selecting on
the ecosystem-scale CH~4~ oxidation rate.  Using this approach, we were able to
estimate how much of the variation in CH~4~ oxidation rate was associated with
variation in the microbiome.  To identify which microorganisms could be markers
of CH~4~ oxidation rate, we compared the composition of the artificially
selected microbiomes to a control set of microbiomes with random selection.  We
then evaluated whether these markers met our underlying assumptions about
which taxa limit the rate of ecosystem function.

@@

Methane oxidation rates were determined after flushing and spiking jars to
approximately 1000 ppm CH~4~. Headspace samples of 1 mL were collected from
each jar immediately after spiking and then at time points 3, 6, 24, and 48
hours for a 5-point curve. Samples were immediately injected into a SRI model
8610C gas chromatograph equipped with a flame ionization detector (SRI
Instruments, Torrance, CA, USA) to determine the headspace CH~4~ concentration.
Fluxes were calculated from a first-order exponential decay function as decay
constant $k$ with units day^-1^. Oxidation rates are presented as the additive
inverse of $k$ (i.e., $-k$) so that a more positive value represents a greater
oxidation rate. The jars selected for the positive treatment in passage 2 had
the lowest CH~4~ oxidation rate of the twelve jars due to a calculation error
in the CH~4~ oxidation rate. Despite this error, we still observed a strong
response to selection demonstrating the power of this technique. In addition,
this likely reintroduced diversity lost during the selection process,
potentially increasing the variation available to be selected upon, as
is recommended for microbiome selection experiments [@sanchez2021].

@@

Bioinformatics processing was performed in 'R' [@rcoreteam2018].
Demultiplexed sequencing reads were denoised using 'DADA2' to generate a table
of amplicon sequence variants (ASVs)
[@callahan2016]. Taxonomic assignment was performed
using the Ribosomal Database Project naive Bayesian classifier [@wang2007]. The presence of
contaminants was evaluated using both the prevalence and frequency methods from
'DECONTAM' [@davis2018].
Decontam identified `r length(which(contamdf$contaminant))` potential contaminants based
on prevalence and frequency. Visual inspection of abundance-concentration
plots indicated that `r length(contaminants)` of these were likely contaminants and these
ASVs were removed. Amplicon sequence variants that were assigned chloroplast taxonomy were removed prior to analysis.

@@

To estimate the proportion of variance in CH~4~ oxidation rate due to variation in
the microbiome, we estimated narrow-sense heritability ($h^2$) as the
regression of offspring on mid-parent [@falconer1996]. The mid-parent was the
mean for all three selected jars and the offspring was the CH~4~ oxidation rate for each jar produced by those parents. First, we tested if there was an effect of treatment
on the heritability estimate. We compared nested models with and without treatment using an F-test
test with the 'anova' function. We then fit the full model to estimate the
heritability of the Positive and Neutral treatments.

@@

To identify taxa that responded to selection on CH~4~ oxidation rate, we tested
differential abundance between the two treatments in passage 5. We first
grouped ASVs at the family level keeping ASVs that lacked a family-level taxonomic
assignment by grouping them at a higher level. We
then subset the samples in Passage 5 and removed all families with a prevalence
of less than 10\% in either treatment. We used three methods for testing differential abundance: ANCOM-II, ALDEx2, and CORNCOB. We then
identified the consensus taxa that were significant
with all three tests and plotted their relative abundances. For ANCOM-II, we used the 'ancom' function in the 'ANCOM-BC' package with a cutoff of W = 0.7 [@kaul2017; @lin2020]. For ALDEx2, we used the 'aldex' function in the 'ALDEx2' package with Welch's t-test and we used an effect size of
1 as our significance threshold [@fernandes2014]. Finally, we used CORNCOB with the 'differentialTest' function in the 'corncob' package with the Wald test and without bootstrapping [@martin2020]. Lastly, to test differentially abundant methanotrophs, we subset all ASVs within methanotrophic families and tested their differential abundance aggregated at the family and genus level using 'corncob'.

@@

We observed a response to selection on whole-ecosystem soil CH~4~ oxidation
rate (`r figures("response", display = "cite")`; difference of slopes:
F~`r compare_response_mods[2,4]`,`r compare_response_mods[2,2]`~ = `r myround(compare_response_mods[2,5], 2)`, p = `r myround(compare_response_mods[2,6], 2)`). At the start of the experiment, the Positive
treatment had a mean CH~4~ oxidation rate that was
`r percent_less(initial = tidy(response_model)[[1, 2]], final = tidy(response_model)[[3, 2]] + tidy(response_model)[[1, 2]])`\% lower than the
Neutral treatment `r print_lm(response_model, 3, 'difference of y-intercepts')`. 
There was no change in CH~4~ oxidation rate of the Neutral
treatment over the five passages `r print_lm(response_model, 2)`. By contrast,
the Positive treatment had a `r round(per_slope)`\% increase in CH~4~
oxidation rate per passage `r print_lm(response_model, 4)`.

We estimated the proportion of variation in methane oxidation due to variation
in the microbiome as the regression of offspring on mid-parent 
(`r figures("herit", display = "cite")`).  Offspring CH~4~ oxidation rates were
correlated with parental CH~4~ oxidation rates in both the Positive treatment
`r print_lm(herit_model, 4)` and the Neutral treatment `r print_lm(herit_model,
2)`.

@@

# Discussion

While it is well understood that the microbiome mediates a variety of ecosystem
functions, a fundamental question in microbial ecology is whether we need to
understand variation in the composition of microbiomes to understand variation
in the rate of ecosystem functions [@conrad1996; @schimel1998]. To address this question, we performed
artificial ecosystem selection on soil microbiomes by selecting for ecosystems
with high methane oxidation rates. We found that there was a significant
increase in methane oxidation rate as a response to selection. Each generation
of soil ecosystems was created from the same pool of sterile soil substrate and
a small inoculum from the previous generation. Therefore, it is unlikely that
this response to selection was due to changes in the abiotic conditions of the
ecosystem. In addition, the response was only observed in the Positive
treatment and not in the Neutral treatment, which shows that this was not a
side effect of passaging in jars.  Therefore, we conclude that variation in the
microbiome can generate variation in ecosystem functions independent of the
environment.

In quantitative genetics, heritability quantifies the degree to which variation
in a trait is due to variation in genotype [@falconer1996]. This is often estimated as the
regression of offspring phenotype on parental phenotype. Thus, a significant
correlation between the trait values of parents and offspring indicates that
the trait is at least partially genetically determined. To estimate the
variation in an ecosystem trait due to variation in the microbiome, we can
perform a similar estimation by comparing the trait values of the "offspring"
jars in one generation to the trait value of the "parental" jars chosen for
selection in the previous generation. This "ecosystem heritability" then
quantifies how much of the variation in ecosystem function is due to variation
in the microbiome [@goodnight2000; @sanchez2021; @henry2021].

We quantified how much of the variation in methane oxidation
rate was due to variation in the microbiome by regressing the methane oxidation
rates of the 12 offspring jars against the mean methane oxidation rate of the
three jars selected in the previous generation 
We observed a heritability of `r myround(coefficients(herit_model)[4], 2)` (add SE/CI) for the
Positive selection treatment. 
A heritability of 1 would indicate that a trait is entirely
genetically determined and thus offspring would have the exact same value of
that trait as their parents. A heritability greater than 1 indicates that
offspring exceed the value of their parents. This is unlikely in the context of
a genomic trait, but microbiomes are not limited by the constraints of genomes
because microbes are not alleles and communities are not chromosomes.
Therefore, complementarity within a microbial community could result in
offspring communities that exceed the trait values of their "parent"
ecosystems. The large heritability we observed indicates that microbiome
variation is a large contributor to variation in methane oxidation in this
experimental system.

@@

Richness at the ASV level did not vary between the two treatments and there
were relatively few taxa gained or lost in the Positive selection treatment and
none of these were prevalent across the 12 jars in passage 5. Therefore, the
gain or loss of species is unlikely to explain the increase in methane
oxidation rate. However, we found that Bray-Curtis dissimilarity was greater
between the two treatments in Passage 5 than within each treatment and was
correlated with methane oxidation rate, which suggests that changes in the
relative abundance of taxa could explain the response to selection.

Even though we observed an increase in methane oxidation rate in the Positive
treatment and a difference in composition between the two treatments, we did
not observe an increase in the relative abundance of methanotrophs. This was
surprising given that methane consumption is not a common trait among microbes
and that it is often assumed that the rate of an ecosystem function is
limited by the final enzymatic step in a metabolic pathway
[@rocca2015]. In certain ecosystems, methane production and consumption are 
correlated with the abundance of methanogens and methanotrophs as estimated from marker
genes [@meyer2020, @meyer2020a]. However, it does not appear generally true that the rate of an
ecosystem function is limited by the abundance of the gene that encodes the
final step in that pathway [@rocca2015]. Our results suggest that the rate of methane oxidation can be altered by
non-methanotrophs, perhaps through ecological interactions with methanotrophic
species.

Another important question is whether selection at the ecosystem scale resulted
in within-strain evolution. Using 16S rRNA gene sequencing we are unable to
resolve genomic changes within taxa that may have undergone evolution. 
Therefore, it is possible that the response to selection was driven by changes 
within the genomes of methanotrophs. But given
the slow growth rate of methanotrophs, often requiring weeks to years of
serial batch culturing for isolation, this seems like an unlikely explanation 
[@whittenbury1970; @hirayama2014; @kim2018]. Future
work could analyze an artificial ecosystem selection experiment using
genome-resolved metagenomics to try to detect strain-level changes in the
constituent taxa. However, reconstructing genomes from in situ soil samples
has been difficult to achieve and may require enrichment techniques, particularly for methanotrophs which are a relatively rare component of the soil microbiome [@delmont2015]. 
If these challenges could be overcome for soil metagenomics, we could begin to evaluate whether ecosystem selection results in strain-level evolution or enriches for certain metabolic pathways within the microbiome.

As more is understood about the role of the microbiome in regulating ecosystem
functions and the specific taxa involved, more detailed experiments could
selectively add or remove important taxa or manipulate their relative abundance
to start to estimate the quantitative relationship between specific taxa and
the rate of an ecosystem function. Both our differential abundance analysis and
this reductionist approach to manipulating individual taxa assumes that taxa
act individually on ecosystem function. However, it is much more likely that
the microbiome affects the rate of ecosystem function through the interactions
of multiple taxa. We are unable to evaluate this possibility given our small 
sample size, but future analyses with greater replication should test the
role of multiple taxa simultaneously by, for example, using machine learning
methods [@topcuoglu2020].

A fundamental question in microbial ecology is whether variation in the
microbiome contributes to variation in ecosystem function. We performed
artificial ecosystem selection on soil methane oxidation rate to test the
relationship between microbiome variation and soil methane oxidation rate. We
observed a response to selection on soil methane oxidation rate and found a
high heritability for methane oxidation, which suggests that variation in the
microbiome contributes to variation in soil methane oxidation rate.
Surprisingly, we did not observe an increase in the relative of abundance of
methanotrophs as a response to selection. This suggests that the rate of methane 
oxidation could be altered by
non-methanotrophs. Future research could estimate
heritability under greater environmental variation to understand how much of
the variation in methane oxidation rate is due to variation in the microbiome
in a natural ecosystem. In addition, future selection experiments could use
genome-resolved metagenomics to observe within-strain evolution and test
changes in the functional pathways of the microbiome as a response to selection
to determine what aspects of the microbiome explain this increase in oxidation
rate. This study demonstrates that soil methane oxidation rate can vary greatly
with variation in the microbiome and opens up future research opportunities to
establish a more detailed understanding of the relationship between the
microbiome and ecosystem function.
